Formalizing the Kolmogorov extension Theorem in Lean

### 2023-09-16 Manuscript/proof_outline.tex
remove todo
remove remark about type theory in Isabelle vs Lean
rewrite remark about extension to pseudo-metric spaces

\todo[inline]{\leanline{∀ j : J, α j} or \leanline{(j : J) -> α j} ? Explain somewhere that this is the mathlib notation for the product space}

@@

The main steps in our construction, which were previously missing in {\tt mathlib}, are (more mathematical details are below, and in standard textbooks on Probability Theory, e.g.\ \cite{Klenke2013, Kallenberg2020}):
\begin{enumerate}
\item a formalization of (the set-system of) semi-rings (see
  Definition~\ref{def:semi});
\item a definition of additive contents over semi-rings (see Definition~\ref{def:content});
\item a proof that a single probability measure on a Polish space is inner regular with respect to compact sets (see Definition~\ref{def:innerreg}; in fact, we proved a slight generalization using pseudo-metric spaces, found in Lemma~\ref{L:relcoPol} below); 
\item the classical Carathéodory extension theorem, providing us with a candidate for the measure which we want to construct (see Theorem~\ref{T:masseind}); 
\item the Kolmogorov extension theorem, as based on the previous steps (See Theorem~\ref{T1}).
\end{enumerate}
The Kolmogorov extension theorem has been previously formalized in Isabelle/HOL \cite{Immler2012}. This formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical. In addition, {\color{blue}Isabelle works with simple type theory, while Lean is an implementation of dependent theory.}

@@

We note that we extend the standard assumption that all $\alpha_t$ are
separable, complete metric spaces (or Polish, i.e.\ separable and
metrizable through a complete metric) to cover the case of extended
pseude-metric spaces. Such spaces do not satisfy the frequently used
Hausdorff (or t2) property, i.e.\ there can be $x\neq y$ such that all
open balls around $x$ and $y$ overlap. {\color{blue}We note that this
  generalization was only possible since underlying results in
  \leanline{mathlib} were already provided on the same level of
  generality: More precisely,
  \leanline{isCompact_iff_totallyBounded_isComplete}, which shows that
  a set $A \subseteq \alpha$ is compact iff it is complete and totally
  bounded {\color{blue}(see note~\ref{note:tot})}, requires $\alpha$
  to be a uniform space (recall that every metric space is uniform);
  recall such spaces from Remark~\ref{rem:uniform}, and see
  note~\ref{note:tot} for a definition of total boundedness. Since we
  also require the underlying space(s) to be second countable (see
  also in the proof of Lemma~\ref{L:relcoPol}), we have to make a
  countabiliy assumption, which leads us to (extended) pseudo-metric
  spaces by \leanline{UniformSpace.pseudoMetrizableSpace}; see also
  Remark~\ref{rem:uniform} for some more details.}

>>>>

@@

The main steps in our construction, which were previously missing in {\tt mathlib}, are (more mathematical details are below, and in standard textbooks on Probability Theory, e.g.\ \cite{Klenke2013, Kallenberg2020}):
\begin{enumerate}
\item a formalization of (the set-system of) semi-rings (see
  Definition~\ref{def:semi});
\item a definition of additive contents over semi-rings (see Definition~\ref{def:content});
\item a proof that a single probability measure on a Polish space is inner regular with respect to compact sets (see Definition~\ref{def:innerreg}; in fact, we proved a slight generalization using pseudo-metric spaces, found in Lemma~\ref{L:relcoPol} below); 
\item the classical Carathéodory extension theorem, providing us with a candidate for the measure which we want to construct (see Theorem~\ref{T:masseind}); 
\item the Kolmogorov extension theorem, as based on the previous steps (See Theorem~\ref{T1}).
\end{enumerate}
The Kolmogorov extension theorem has been previously formalized in Isabelle/HOL \cite{Immler2012}. This formalization only works on Polish spaces (rather than on spaces where every finite measure is inner regular with respect to compact sets, see below), and only in the case where all $\alpha_i$'s are identical. %In addition, Isabelle/HOL works with simple type theory, while Lean is an implementation of dependent type theory.

@@

We note that we extend the standard assumption that all $\alpha_t$ are
separable, complete metric spaces (or Polish, i.e.\ separable and
metrizable through a complete metric) to cover the case of extended
pseude-metric spaces. Such spaces do not satisfy the frequently used
Hausdorff (or t2) property, i.e.\ there can be $x\neq y$ such that all
open balls around $x$ and $y$ overlap.
  This generalization was only possible since underlying results in
  \leanline{mathlib} were already provided on the same level of
  generality.
  More precisely, \leanline{isCompact_iff_totallyBounded_isComplete}, which shows that
  a set $A \subseteq \alpha$ is compact iff it is complete and totally
  bounded (see note~\ref{note:tot}), requires $\alpha$
  to be a uniform space (Remark~\ref{rem:uniform}, recall that every metric space is uniform).
  We also require the underlying space(s) to be second-countable (used in the proof of Lemma~\ref{L:relcoPol}).
  A second-countable uniform space can be made into an (extended) pseudo-metric space
  (\leanline{UniformSpace.pseudoMetrizableSpace}); see also
  Remark~\ref{rem:uniform} for some more details.
